Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-07 manuscript.Rmd
knit
cache results

```{r participants}
ppts <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")))
num_younger <- ppts %>% filter(group == "Younger") %>% .$id %>% unique(.) %>% length(.)
num_older <-  ppts %>% filter(group == "Older") %>% .$id %>% unique(.) %>% length(.)
num_sessions <- ppts %>% group_by(group, id, session) %>% slice_head(n = 1) %>% count %>% group_by(group, id) %>% count
age <- ppts %>% group_by(group) %>% get_summary_stats(age) %>% column_to_rownames("group")
```

@@

A primary coder completed annotation for the full length of the video, while an independent reliability coder completed annotation for the first thirty minutes of each video. Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code. Overall agreement averaged XX.X% across video files, ranging from XX-XX. Cohen's *kappa* ranged from XX-XX across files, with a mean of XX.

## Body position classification

@@

```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=5, fig.height=8}
metrics <- read_csv("data/group_split_metrics.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

```{r metricstable}

@@

```{r metricsbyclass}
metrics_class <- read_csv("data/group_split_metrics_class.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

```{r part2overall, warning=FALSE, fig.cap="Overall agreement between human-coded body position and model-predicted body position in the distal comparison. Agreement for group models is shown in (A) and agreement for individual models is shown in (B). Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=7, fig.height=6.5}

@@

```{r, pt2overalltable}
#By posture
pt2corr_with_outliers <- map_dfc(agree, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
pt2corr_without_outliers <- map_dfc(agree, ~ .x %>% filter_outliers %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
pt2corr <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
rm(pt2corr_with_outliers)
rm(pt2corr_without_outliers)

@@

```{r part2bins, warning=FALSE, fig.cap="Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5}
calculate_agreement_bins <- function(ds) {
  ds <- ds %>% group_by(file) %>% 
    mutate(sample = row_number(), bin = floor(sample/300)) %>% 
    add_count(file, bin) %>% ungroup() %>% filter(n > 200) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
  model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
  human <- ds %>% count(file, bin, num_bins,  code) %>% add_count(file, bin, num_bins, wt = n, name = "total") %>% mutate(prop_human = n/total)
  agree <- full_join(select(model, file, pos, prop_model, bin, num_bins), 
                     select(human, file, code, prop_human, bin, num_bins), 
                     by = c("file" = "file", "pos" = "code", "bin" = "bin", "num_bins" = "num_bins"))
  agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
                            prop_model = ifelse(is.na(prop_model), 0, prop_model),
                            pos - factor(pos, levels = pos_levels))
  agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
}

@@

```{r pt2binstable}
#Bins
pt2corr_with_outliers <- map_dfc(agree_bins, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
pt2corr_without_outliers <- map_dfc(agree_bins, ~ .x %>% filter_outliers %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
pt2corr_bins <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
rm(pt2corr_with_outliers)
rm(pt2corr_without_outliers)

@@

```{r timelines, warning=FALSE, fig.cap="Timelines", fig.align="center", fig.width=7.75, fig.height=8}
timeline_data <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401) %>% 
  pivot_longer(cols = nap_time:upright_time, names_to = "position", values_to = "prop") %>% 
  mutate(Position = factor(position,
                           levels = c("nap_time","upright_time", "sit_time", "prone_time", "supine_time", "held_time"),
                           labels = c("Nap","Upright", "Sitting", "Prone", "Supine", "Held"))) %>% 
  arrange(Position, age)

@@

```{r age, warning=FALSE, fig.cap="Age trends", fig.align="center", fig.width=7.5, fig.height=7}
fullday_group <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401)
fullday_split <- read_csv("data/imu-compiled-split.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401)
fullday <- list("group" = fullday_group, "split" = fullday_split)

@@

```{r agetable}
pos_sum_age <- map(fullday, fullday_long) %>% 
  map(~ group_by(.x, group, position, id_uni) %>% 
            mutate(prop = prop * 100) %>% 
            get_summary_stats(prop) %>% 
            group_by(group, position) %>% 
            get_summary_stats(mean) %>% 
            select(group, position, mean, sd))
pos_sum_age <- pos_sum_age %>% map2_dfr(c("Group", "Individual"), 
                         ~ mutate(.x, Model = .y,
                                  mean_sd = str_glue("{apa_num(mean, digits = 1)}% ({apa_num(sd, digits = 1)})")) %>% 
                           select(-mean, -sd) %>% 
                           pivot_wider(names_from = group, values_from = mean_sd)) %>% 
  pivot_wider(names_from = Model, values_from = c("Younger", "Older")) %>% 
  relocate(Younger_Individual, .after = Older_Group)
apa_table(pos_sum_age, col.names = c("Position", "Younger", "Older", "Younger", "Older"), col_spanners = list("Group" = c(2, 3), "Individual" = c(4, 5)),
          caption = "Summary of age differences in full-day body position for younger (4- to 7-month) and older (11- to 14-month) infants. Values shown are the mean percent of time for each body position averaged across infants in each group. Standard deviations are shown in parentheses. Descriptive statistics are shown separately for group and individual models.")

>>>>

```{r participants, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
ppts <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")))
num_younger <- ppts %>% filter(group == "Younger") %>% .$id %>% unique(.) %>% length(.)
num_older <-  ppts %>% filter(group == "Older") %>% .$id %>% unique(.) %>% length(.)
num_sessions <- ppts %>% group_by(group, id, session) %>% slice_head(n = 1) %>% count %>% group_by(group, id) %>% count
age <- ppts %>% group_by(group) %>% get_summary_stats(age) %>% column_to_rownames("group")
```

@@

A primary coder completed annotation for the full length of the video, while an independent reliability coder completed annotation for the first thirty minutes of each video. Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code. Overall agreement averaged XX.X% across video files, ranging from XX-XX. Cohen's *kappa* averaged XX.X% across video files, ranging from XX-XX.

## Body Position Classification

@@

```{r metrics, warning=FALSE, fig.cap="Metrics", fig.align="center", fig.width=5, fig.height=8, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}
metrics <- read_csv("data/group_split_metrics.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

```{r metricstable, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}

@@

```{r metricsbyclass, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics_class.csv")}
metrics_class <- read_csv("data/group_split_metrics_class.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

```{r part2overall, warning=FALSE, fig.cap="Overall agreement between human-coded body position and model-predicted body position in the distal comparison. Agreement for group models is shown in (A) and agreement for individual models is shown in (B). Plots are shown separately for each body position with a reference line that indicates perfect agreement; each point in a plot represent data for a single participant. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=7, fig.height=6.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}

@@

```{r, pt2overalltable, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
#By posture
pt2corr_with_outliers <- map_dfc(agree, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
pt2corr_without_outliers <- map_dfc(agree, ~ .x %>% filter_outliers %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
pt2corr <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
rm(pt2corr_with_outliers)
rm(pt2corr_without_outliers)

@@

```{r part2bins, warning=FALSE, fig.cap="Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
calculate_agreement_bins <- function(ds) {
  ds <- ds %>% group_by(file) %>% 
    mutate(sample = row_number(), bin = floor(sample/300)) %>% 
    add_count(file, bin) %>% ungroup() %>% filter(n > 200) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
  model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
  human <- ds %>% count(file, bin, num_bins,  code) %>% add_count(file, bin, num_bins, wt = n, name = "total") %>% mutate(prop_human = n/total)
  agree <- full_join(select(model, file, pos, prop_model, bin, num_bins), 
                     select(human, file, code, prop_human, bin, num_bins), 
                     by = c("file" = "file", "pos" = "code", "bin" = "bin", "num_bins" = "num_bins"))
  agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
                            prop_model = ifelse(is.na(prop_model), 0, prop_model),
                            pos - factor(pos, levels = pos_levels))
  agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
}

@@

```{r pt2binstable, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
#Bins
pt2corr_with_outliers <- map_dfc(agree_bins, ~ .x %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...3`)
pt2corr_without_outliers <- map_dfc(agree_bins, ~ .x %>% filter_outliers %>% group_by(pos) %>% cor_test(vars = c("prop_model", "prop_human")) %>% select(pos, cor)) %>% select(-`pos...1`,-`pos...3`)
pt2corr_bins <- bind_cols(pt2corr_with_outliers, pt2corr_without_outliers)
rm(pt2corr_with_outliers)
rm(pt2corr_without_outliers)

@@

```{r timelines, warning=FALSE, fig.cap="Timelines", fig.align="center", fig.width=7.75, fig.height=8, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
timeline_data <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401) %>% 
  pivot_longer(cols = nap_time:upright_time, names_to = "position", values_to = "prop") %>% 
  mutate(Position = factor(position,
                           levels = c("nap_time","upright_time", "sit_time", "prone_time", "supine_time", "held_time"),
                           labels = c("Nap","Upright", "Sitting", "Prone", "Supine", "Held"))) %>% 
  arrange(Position, age)

@@

```{r age, warning=FALSE, fig.cap="Age trends", fig.align="center", fig.width=7.5, fig.height=7, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
fullday_group <- read_csv("data/imu-compiled.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401)
fullday_split <- read_csv("data/imu-compiled-split.csv") %>% 
  mutate(id_uni = factor(id*100+session),
         group = factor(ifelse(age <= 8, "Younger", "Older")),
         nap_time = ifelse(is.na(sit_time), nap_time, NA)) %>% 
  filter(id_uni != 10401)
fullday <- list("group" = fullday_group, "split" = fullday_split)

@@

```{r agetable, cache=TRUE, cache.extra=tools::md5sum("data/imu-compiled.csv")}
pos_sum_age <- map(fullday, fullday_long) %>% 
  map(~ group_by(.x, group, position, id_uni) %>% 
            mutate(prop = prop * 100) %>% 
            get_summary_stats(prop) %>% 
            group_by(group, position) %>% 
            get_summary_stats(mean) %>% 
            select(group, position, mean, sd))
pos_sum_age <- pos_sum_age %>% map2_dfr(c("Group", "Individual"), 
                         ~ mutate(.x, Model = .y,
                                  mean_sd = str_glue("{apa_num(mean, digits = 1)}% ({apa_num(sd, digits = 1)})")) %>% 
                           select(-mean, -sd) %>% 
                           pivot_wider(names_from = group, values_from = mean_sd)) %>% 
  pivot_wider(names_from = Model, values_from = c("Younger", "Older")) %>% 
  relocate(Younger_Individual, .after = Older_Group)
apa_table(pos_sum_age, col.names = c("Position", "Younger", "Older", "Younger", "Older"), col_spanners = list("Group" = c(2, 3), "Individual" = c(4, 5)),
          caption = "Summary of age differences in full-day body position for younger (4- to 7-month) and older (11- to 14-month) infants. Values shown are the mean percent of time for each body position averaged across infants in each group. Standard deviations are shown in parentheses. Descriptive statistics are shown separately for group and individual models.")

### 2023-03-07 manuscript.tex
cache results

A primary coder completed annotation for the full length of the video, while an independent reliability coder completed annotation for the first thirty minutes of each video. Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code. Overall agreement averaged XX.X\% across video files, ranging from XX-XX. Cohen's \emph{kappa} ranged from XX-XX across files, with a mean of XX.

\hypertarget{body-position-classification}{%
\subsection{Body position classification}\label{body-position-classification}}

>>>>

A primary coder completed annotation for the full length of the video, while an independent reliability coder completed annotation for the first thirty minutes of each video. Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code. Overall agreement averaged XX.X\% across video files, ranging from XX-XX. Cohen's \emph{kappa} averaged XX.X\% across video files, ranging from XX-XX.

\hypertarget{body-position-classification}{%
\subsection{Body Position Classification}\label{body-position-classification}}
